---
title: "Counted hierarchies around numeric radar profiles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Counted hierarchies around numeric radar profiles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 10,
                      fig.height = 7, dpi = 96)
library(radarsunburst)
```

## The two questions answered by this chart

The centre compares arithmetic means of numeric cells within groups. The outside
counts records along a hierarchy. These are different summaries of the same
table: neither the groups nor metric names need to match the root categories.
The package does not infer a causal relationship between the centre and outside.

```{r data}
data(radar_sunburst_demo)
d <- radar_sunburst_demo
metrics <- c("score_stability", "score_efficiency", "score_innovation",
             "score_quality", "score_collaboration")
hierarchy <- c("main_category", "sub_category", "leaf_category")
s <- radar_sunburst_data(d, metrics, "series", hierarchy)
s$radar
s$nodes[s$nodes$level == 1,
        c("label", "count", "root_proportion", "global_proportion")]
```

Every row counts once. A weighting column, ID or project name is not required.
If a source has duplicate rows, deduplicate it explicitly before calling the
package when that is the intended scientific counting unit. A table with one
row per group uses those numeric values unchanged. Blank child cells terminate
a branch; their angular share stays unpainted in outer levels rather than being
redistributed to the remaining children.

```{r chart}
p <- radar_sunburst(d, metrics, "series", hierarchy,
  metric_labels = c("Stability", "Efficiency", "Innovation", "Quality",
                    "Collaboration"),
  root_order = c("F", "C", "D", "B", "E"),
  root_colors = c(B = "#FF954F", C = "#F45668", D = "#DE578A",
                   E = "#AE4D7C", F = "#725993"),
  group_colors = c("#8E176C", "#1485C8", "#00A58E", "#C78800"))
p
```

## Counts, percentages and ordering

The root D has eight records, its child D1 four, and D1-1 two; the full table has
40. Thus their root-based percentages are 100%, 50%, 25%; global percentages are
20%, 10%, 5%; parent percentages are 20%, 50%, 50%. Top-level categories each
showing 100% under the root convention should not be added together.

```{r proportions}
selected <- s$nodes$label %in% c("D", "D1", "D1-1")
s$nodes[selected, c("label", "count", "parent_count", "root_count",
                    "parent_proportion", "root_proportion", "global_proportion")]
count_only <- radar_sunburst(d, metrics, "series", hierarchy, percent = "none")
```

Use `percent = "parent"`, `"root"`, `"global"` or `"none"`. This controls labels
and hover denominators, not the underlying counts or sector widths. Defaults
preserve first appearance. Supply complete `root_order` and `group_order`
vectors to reorder them. For a local child order, obtain the parent's ID from
the summary, then use a named list; full paths are stored separately as lists
so slash characters or repeated names cannot merge nodes.

```{r ordering}
parent_id <- s$nodes$id[s$nodes$level == 1 & s$nodes$label == "F"]
orders <- stats::setNames(list(c("F2", "F1")), parent_id)
reordered <- radar_sunburst_data(d, metrics, "series", hierarchy,
                                 branch_order = orders)
reordered$nodes[reordered$nodes$parent_id %in% parent_id, c("label", "count")]
```

## Shared radar limits

The default uses all raw input metric values before averaging. The observed
range is 30 to 97, which expands to 26.65 to 100.35. The centre labels 26.65,
not zero. Fixed limits make charts from different subsets easier to compare.
Values outside explicit limits raise an error with the column and row location.
Constant data get a nonzero range; negative and zero values are supported.

```{r limits}
s$radar_limits
fixed <- radar_sunburst(d, metrics, "series", hierarchy, radar_limits = c(0, 100))
attr(fixed, "radar_sunburst_summary")$radar_limits
```

All metrics use the same range. The package does not automatically make budget,
team size or duration comparable to scores; prepare comparable metrics explicitly
when that interpretation is needed.

## Standard themes and scales

```{r themes}
p + ggplot2::theme_minimal() + ggplot2::labs(title = "Fictional projects")
```

`theme_classic()`, `theme_bw()` and `theme_dark()` also work. A third-party theme
is optional and used only when installed.

```{r third-party, fig.show='hide'}
if (requireNamespace("ggthemes", quietly = TRUE)) {
  p + ggthemes::theme_few()
}
```

The metric ring is omitted entirely by default (`metric_ring = NULL`). Add it
with `metric_ring = c(1.06, 1.55)`; use `show_metric_labels = TRUE` to label it.
Radar hover always retains the metric names supplied in `metric_labels`.
Theme components control titles, guides and backgrounds. Geometry labels have
separate `label_size`, `metric_label_size`, `min_label_size`, `label_family`,
`label_colour`, `metric_label_colour`, and `label_angle` arguments. Dark or unusual
backgrounds can require explicit label colours. Check font availability on every
target platform. Coloured metric sectors do not imply metric-category matching.

```{r palettes, fig.show='hide'}
p + ggplot2::scale_colour_manual(values = c(Series1 = "#444444", Series2 = "#0072B2",
                                           Series3 = "#009E73", Series4 = "#D55E00"))
p + ggplot2::theme(legend.position = "bottom")
```

Label fitting uses conservative bounds for a typical 180 mm plot panel.
Very dense branches or long text are reduced and then hidden below the minimum
size; the underlying sectors and hover information remain. Export size and font
metrics affect readability, so inspect the final artifact. Custom ring radii and
label sizes allow a different layout without changing the numerical summaries.

## Static and interactive export

The following examples write only to temporary directories and clean up.
For your own persistent files, explicitly choose an output path.

```{r static-export}
local({
  directory <- tempfile("radarsunburst-static-")
  dir.create(directory)
  on.exit(unlink(directory, recursive = TRUE), add = TRUE)
  ggplot2::ggsave(file.path(directory, "chart.png"), p,
                  width = 12, height = 8, dpi = 150)
  if (capabilities("cairo")) {
    ggplot2::ggsave(file.path(directory, "chart.svg"), p,
                    width = 12, height = 8, device = grDevices::svg)
  }
})
```

```{r interactive, eval=TRUE}
if (requireNamespace("plotly", quietly = TRUE)) {
  w <- radar_sunburstly(p, width = 800, height = 550)
  # Direct standard conversion is also supported.
  direct <- plotly::ggplotly(p, tooltip = "text")
  inherits(w, "plotly")
}
```

Hover shows the full path, count, denominator and percentage, or the radar group,
metric and mean. Zoom preserves the circle's aspect ratio. A radar legend click
toggles its points, line and fill together; category legends remain explanatory.
Decorative layers are excluded from hover by the configured converter. There is
no branch drill-down or linked filtering. Geometry text rotation and some custom
theme extensions may differ after conversion; horizontal text is the default.

```{r html-export}
if (requireNamespace("plotly", quietly = TRUE) &&
    requireNamespace("htmlwidgets", quietly = TRUE)) {
  local({
    directory <- tempfile("radarsunburst-html-")
    dir.create(directory)
    on.exit(unlink(directory, recursive = TRUE), add = TRUE)
    htmlwidgets::saveWidget(radar_sunburstly(p), file.path(directory, "chart.html"),
                            selfcontained = FALSE)
  })
}
```

`selfcontained = FALSE` saves the JavaScript assets beside the HTML, without a
network request when the saved result is opened. A self-contained export requires
Pandoc in the author's development environment, not for ordinary plotting.
